---
title: "DATS6101 Group Project: Predicting Strokes"
author: "Jonathan Giguere, Jesse Borg, Ese Emuraye, Sarah Gates"
date: "December 11th 2019"
output: 
   html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

Start by loading the packages we know will be needed for the project. Then load the brfss 2013 data which is as an RData file as opposed to a csv.

```{r load-packages, message = FALSE}
loadPkg('ggplot2')
loadPkg('dplyr')
loadPkg('plyr')
```

```{r load-data}
load("brfss2013.RData")
```

### Cleaning the data

We started by choosing the variables which we think will be useful or the most interesting for this study and then renamed them to have more logical or descriptive names. 
```{r keeps}
# Choose which variables we want to keep

keeps <- c('X_state', 'imonth', 'genhlth', 'menthlth', 'hlthpln1', 'sleptim1', 'bphigh4', 'bpmeds', 'cholchk', 'cvdinfr4', 'cvdcrhd4', 'cvdstrk3', 'asthma3', 'asthnow', 'addepev2', 'chckidny', 'diabete3', 'veteran3', 'marital', 'educa', 'employ1', 'income2', 'weight2', 'sex', 'pregnant', 'diffwalk', 'smoke100', 'smokday2', 'stopsmk2', 'usenow3', 'alcday5', 'avedrnk2', 'drnk3ge5', 'fruit1', 'fvgreen', 'exerany2', 'prediab1', 'qlmentl2', 'qlstres2', 'drvisits', 'ssbsugar', 'cvdasprn', 'scntwrk1','X_ageg5yr')

brfss_reduced <- brfss2013[keeps]
```

```{r names}
# Rename the variables to be more descriptive

new_names <- c('state', 'month', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'bp_meds', 'time_since_cholcheck', 'heart_attack', 'angina', 'stroke', 'asthma', 'has_asthma_now', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'pregnant', 'difficulty_walk', 'smoke_100', 'freq_smoke', 'stop_smoke_year', 'smokeless_tabac', 'alc_past_30', 'alc_perday_30', 'binge_alc', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'prediabetes', 'depressed_30', 'anxious_30', 'dr_visits_year', 'soda_freq', 'aspirin_daily', 'work_hours_week','age5yr_bucket')

names(brfss_reduced) <- new_names
```

Once this was done, check to see how many missing values for each column.
```{r check for NAs}
missing_vals <- colSums(is.na(brfss_reduced))
sort(missing_vals)
```

Since there are some columns with a lot of missing values, we decided to drop the ones with more than 100,000 missing entries.
```{r}
# Drop the variables which have too many missing values

brfss_reduced_drops <- subset(brfss_reduced, select = -c(dr_visits_year, prediabetes, binge_alc, alc_perday_30, freq_smoke, bp_meds, aspirin_daily, soda_freq, pregnant, stop_smoke_year, has_asthma_now, work_hours_week, depressed_30, anxious_30))
```

We then removed the rows which have any missing values so that the data we have is complete. If there are enough rows left, we will keep it like this, if not we will have to use rows with missing values.

```{r}
# Remove the rows which have na values in any of the remaining variables

brfss_complete <- na.omit(brfss_reduced_drops)
```

Inspect the dataframe after the missing values have been removed, which shows that there are still over 300,000 observations and 30 variables. This is more than enough data for what we need. 
```{r}
# Inspect the dataframe

str(brfss_complete)
```

For the factors, we want to change some of the variable factors to be more intuitive/descriptive, start by inspecting the levels for the factors we might want to change.

```{r}
# Check the levels for some varibles

levels(brfss_complete$high_bp)
levels(brfss_complete$employment_status)
levels(brfss_complete$smokeless_tabac)
levels(brfss_complete$diabetes)
```

Change the yes/no factors to 1/0 for model purposes and change other factors in the variables above to make them better grouped.

```{r}
# Change the factor levels for the variables wwe feel would be better off changed

levels(brfss_complete$health_coverage)[levels(brfss_complete$health_coverage) == 'Yes'] <- 1
levels(brfss_complete$health_coverage)[levels(brfss_complete$health_coverage) == 'No'] <- 0

levels(brfss_complete$heart_attack)[levels(brfss_complete$heart_attack) == 'Yes'] <- 1
levels(brfss_complete$heart_attack)[levels(brfss_complete$heart_attack) == 'No'] <- 0

levels(brfss_complete$angina)[levels(brfss_complete$angina) == 'Yes'] <- 1
levels(brfss_complete$angina)[levels(brfss_complete$angina) == 'No'] <- 0

levels(brfss_complete$stroke)[levels(brfss_complete$stroke) == 'Yes'] <- 1
levels(brfss_complete$stroke)[levels(brfss_complete$stroke) == 'No'] <- 0

levels(brfss_complete$asthma)[levels(brfss_complete$asthma) == 'Yes'] <- 1
levels(brfss_complete$asthma)[levels(brfss_complete$asthma) == 'No'] <- 0

levels(brfss_complete$depression)[levels(brfss_complete$depression) == 'Yes'] <- 1
levels(brfss_complete$depression)[levels(brfss_complete$depression) == 'No'] <- 0

levels(brfss_complete$kidney_disease)[levels(brfss_complete$kidney_disease) == 'Yes'] <- 1
levels(brfss_complete$kidney_disease)[levels(brfss_complete$kidney_disease) == 'No'] <- 0

levels(brfss_complete$veteran)[levels(brfss_complete$veteran) == 'Yes'] <- 1
levels(brfss_complete$veteran)[levels(brfss_complete$veteran) == 'No'] <- 0

levels(brfss_complete$difficulty_walk)[levels(brfss_complete$difficulty_walk) == 'Yes'] <- 1
levels(brfss_complete$difficulty_walk)[levels(brfss_complete$difficulty_walk) == 'No'] <- 0

levels(brfss_complete$smoke_100)[levels(brfss_complete$smoke_100) == 'Yes'] <- 1
levels(brfss_complete$smoke_100)[levels(brfss_complete$smoke_100) == 'No'] <- 0

levels(brfss_complete$exercise_30)[levels(brfss_complete$exercise_30) == 'Yes'] <- 1
levels(brfss_complete$exercise_30)[levels(brfss_complete$exercise_30) == 'No'] <- 0

levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == "Yes, but female told only during pregnancy"] <- 1
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == "No, pre-diabetes or borderline diabetes"  ] <- 0
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == 'Yes'] <- 1
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == 'No'] <- 0

levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == "Yes, but female told only during pregnancy"] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == "Told borderline or pre-hypertensive" ] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == 'Yes'] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == 'No'] <- 0

levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Employed for wages" ] <- 'Employed'
levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Out of work for 1 year or more" ] <- 'Unemployed'
levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Out of work for less than 1 year" ] <- 'Unemployed'

levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Every day" ] <- 1
levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Some days" ] <- 1
levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Not at all" ] <- 0

```

Inspect the structure of the dataframe with changed variables

```{r}
str(brfss_complete)
```


### EDA

How many participants have smoked at least 100 cigarettes?

```{r}
ggplot(brfss_complete, aes(x = smoke_100, fill = sex)) +
        geom_bar() + 
        ggtitle('Number of participants to have smoked at least 100 cigarettes') +
        ylab('Total') +
        xlab('') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

How many participants use smokeless tobacco products?

```{r}
ggplot(brfss_complete, aes(x = smokeless_tabac, fill = sex)) +
        geom_bar() + 
        ggtitle('Number of participants to have used smokeless tobacco products') +
        ylab('Total') +
        xlab('') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Alcohol

```{r}
hist(brfss_complete$alc_past_30,
     main = 'Frequency of alcohol consumption in 2013 among participants',
     xlab = 'Alcohol Consumption',
     col = magma(3),
     breaks=seq(0,250,by=50))
```

How many times did participants have fruit in the past year?

```{r}
hist(brfss_complete$fruit_freq,
     main = 'Frequency of fruit consumption in 2013 among participants',
     xlab = 'Fruit Consumption',
     col = viridis(7),
     breaks=seq(0,400,by=50))
```

How many times did participants eat dark green vegetables in the past year?

```{r}
hist(brfss_complete$green_veg_freq,
     main = 'Frequency of dark green vegetable consumption in 2013 among participants',
     xlab = 'Dark Green Vegetable Consumption',
     col = cividis(3))
```

How many participants have exercised in the past 30 days

```{r}
ggplot(brfss_complete, aes(x = exercise_30, fill = age5yr_bucket)) +
        geom_bar() + 
        ggtitle('Number of participants to have exercised in the past 30 days') +
        ylab('Total') +
        xlab('') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
        scale_fill_manual(values=plasma(13))
```