---
title: "DATS6101 Group Project: Predicting Heart Attacks"
author: "Jonathan Giguere, Jesse Borg, Ese Emuraye, Sarah Gates"
date: "December 11th 2019"
output: 
   html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

## Loading the Data

Start by loading the packages we know will be needed for the project. Then load the brfss 2013 data which is as an RData file as opposed to a csv.

```{r load-packages, message = FALSE, echo=TRUE}
loadPkg('ggplot2')
loadPkg('dplyr')
loadPkg('usmap')
loadPkg('viridis')
```

```{r load-data, echo=TRUE}
load("brfss2013.RData")
```

## Cleaning the data

We started by choosing the variables which we think will be useful or the most interesting for this study and then renamed them to have more logical or descriptive names.

```{r keeps, echo=TRUE}
# Choose which variables we want to keep

keeps <- c('X_state', 'imonth', 'genhlth', 'menthlth', 'hlthpln1', 'sleptim1', 'bphigh4', 'bpmeds', 'cholchk','toldhi2', 'cvdinfr4', 'cvdcrhd4', 'cvdstrk3', 'asthma3', 'asthnow', 'addepev2', 'chckidny', 'diabete3', 'veteran3', 'marital', 'educa', 'employ1', 'income2', 'weight2', 'sex', 'pregnant', 'diffwalk', 'smoke100', 'smokday2', 'stopsmk2', 'usenow3', 'alcday5', 'avedrnk2', 'drnk3ge5', 'fruit1', 'fvgreen', 'exerany2', 'prediab1', 'qlmentl2', 'qlstres2', 'drvisits', 'ssbsugar', 'cvdasprn', 'scntwrk1','X_ageg5yr')

brfss_reduced <- brfss2013[keeps]
```


```{r rename, echo=TRUE}
# Rename the variables to be more descriptive


new_names <- c('state', 'month', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'bp_meds', 'time_since_cholcheck','told_high_chol', 'heart_attack', 'angina', 'stroke', 'asthma', 'has_asthma_now', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'pregnant', 'difficulty_walk', 'smoke_100', 'freq_smoke', 'stop_smoke_year', 'smokeless_tabac', 'alc_past_30', 'alc_perday_30', 'binge_alc', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'prediabetes', 'depressed_30', 'anxious_30', 'dr_visits_year', 'soda_freq', 'aspirin_daily', 'work_hours_week','age5yr_bucket')

names(brfss_reduced) <- new_names
```

Once this was done, check to see how many missing values for each column.
```{r check for NAs}
missing_vals <- colSums(is.na(brfss_reduced))
sort(missing_vals)
```

Since there are some columns with a lot of missing values, we decided to drop the ones with more than 100,000 missing entries.
```{r, echo=TRUE}
# Drop the variables which have too many missing values

brfss_reduced_drops <- subset(brfss_reduced, select = -c(dr_visits_year, prediabetes, binge_alc, alc_perday_30, freq_smoke, bp_meds, aspirin_daily, soda_freq, pregnant, stop_smoke_year, has_asthma_now, work_hours_week, depressed_30, anxious_30))
```

We then removed the rows which have any missing values so that the data we have is complete. 

```{r, echo=TRUE}
# Remove the rows which have na values in any of the remaining variables

brfss_complete <- na.omit(brfss_reduced_drops)
```

Inspect the dataframe after the missing values have been removed, which shows that there are still over 300,000 observations and 31 variables. This is more than enough data for what we need. 
```{r}
# Inspect the dataframe

str(brfss_complete)
```

For the factors, we want to change some of the variable factors to be more intuitive/descriptive, start by inspecting the levels for the factors we might want to change.

```{r, echo=TRUE}
# Check the levels for some varibles

levels(brfss_complete$high_bp)
levels(brfss_complete$employment_status)
levels(brfss_complete$smokeless_tabac)
levels(brfss_complete$diabetes)
```

Change the yes/no factors to 1/0 for model purposes and change other factors in the variables above to make them better grouped.

```{r, echo=TRUE}
# Change the factor levels for the variables wwe feel would be better off changed

levels(brfss_complete$health_coverage)[levels(brfss_complete$health_coverage) == 'Yes'] <- 1
levels(brfss_complete$health_coverage)[levels(brfss_complete$health_coverage) == 'No'] <- 0

levels(brfss_complete$heart_attack)[levels(brfss_complete$heart_attack) == 'Yes'] <- 1
levels(brfss_complete$heart_attack)[levels(brfss_complete$heart_attack) == 'No'] <- 0

levels(brfss_complete$angina)[levels(brfss_complete$angina) == 'Yes'] <- 1
levels(brfss_complete$angina)[levels(brfss_complete$angina) == 'No'] <- 0

levels(brfss_complete$stroke)[levels(brfss_complete$stroke) == 'Yes'] <- 1
levels(brfss_complete$stroke)[levels(brfss_complete$stroke) == 'No'] <- 0

levels(brfss_complete$asthma)[levels(brfss_complete$asthma) == 'Yes'] <- 1
levels(brfss_complete$asthma)[levels(brfss_complete$asthma) == 'No'] <- 0

levels(brfss_complete$depression)[levels(brfss_complete$depression) == 'Yes'] <- 1
levels(brfss_complete$depression)[levels(brfss_complete$depression) == 'No'] <- 0

levels(brfss_complete$kidney_disease)[levels(brfss_complete$kidney_disease) == 'Yes'] <- 1
levels(brfss_complete$kidney_disease)[levels(brfss_complete$kidney_disease) == 'No'] <- 0

levels(brfss_complete$veteran)[levels(brfss_complete$veteran) == 'Yes'] <- 1
levels(brfss_complete$veteran)[levels(brfss_complete$veteran) == 'No'] <- 0

levels(brfss_complete$difficulty_walk)[levels(brfss_complete$difficulty_walk) == 'Yes'] <- 1
levels(brfss_complete$difficulty_walk)[levels(brfss_complete$difficulty_walk) == 'No'] <- 0

levels(brfss_complete$smoke_100)[levels(brfss_complete$smoke_100) == 'Yes'] <- 1
levels(brfss_complete$smoke_100)[levels(brfss_complete$smoke_100) == 'No'] <- 0

levels(brfss_complete$exercise_30)[levels(brfss_complete$exercise_30) == 'Yes'] <- 1
levels(brfss_complete$exercise_30)[levels(brfss_complete$exercise_30) == 'No'] <- 0

levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == "Yes, but female told only during pregnancy"] <- 1
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == "No, pre-diabetes or borderline diabetes"  ] <- 0
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == 'Yes'] <- 1
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == 'No'] <- 0

levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == "Yes, but female told only during pregnancy"] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == "Told borderline or pre-hypertensive" ] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == 'Yes'] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == 'No'] <- 0

levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Employed for wages" ] <- 'Employed'
levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Out of work for 1 year or more" ] <- 'Unemployed'
levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Out of work for less than 1 year" ] <- 'Unemployed'

levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Every day" ] <- 1
levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Some days" ] <- 1
levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Not at all" ] <- 0

```

Inspect the structure of the dataframe with changed variables

```{r}
str(brfss_complete)
```

## Exploratory Data Analysis

### Jonathan EDA
- Heart attacks and states?
- General  health and effects on heart attack?
- Health coverage and heart attack?
- Sleep and heart attack?
- High BP and heart attack?

```{r states and heart attack data}
#Making dataframe for visualizing heart attacks per state
#Notice that in order to add up the number of heart attacks, i convert the factor to a string and then an integer
state_map_df <- brfss_complete %>% select(state, heart_attack) %>% mutate(state = as.character(state)) %>% group_by(state) %>% summarise(sum_ = sum(as.numeric(as.character(heart_attack)))) %>% filter(state != c('District of Columbia', 'Puerto Rico', 'Guam'))
```

```{r states and heart map}
#Map showing number of heart attacks per
plot_usmap(data = state_map_df, values = "sum_", color = "white") + 
  scale_fill_continuous(name = "Heart Attack Count", label = scales::comma) + 
  labs(title = 'Number of Heart Attacks per State') +
  theme(legend.position = "right")
```

```{r states and participants data}
#Make dataframe showing number of survey particpants by state
participants_map_df <-  brfss_complete %>% group_by(state) %>% summarise(n = n())
```

```{r states and respondents}
#Visualize map of number of participants per state
plot_usmap(data = participants_map_df, values = "n", color = "white") + 
  scale_fill_continuous(name = "Survey Participants by State", label = scales::comma) + 
  labs(title = 'Number of Survey Participants per State') +
  theme(legend.position = "right") 
```

```{r heart attacks for different gen_health levels}
#Visualize proportion of respondents with heart attacks by gen_health
ggplot(brfss_complete, aes(x = gen_health, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Heart attack among levels of general health') +
        ylab('Total') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r heart attacks based on health coverage}
#Visualize proportion of respondents with heart attacks by health_coverage
ggplot(brfss_complete, aes(x = health_coverage, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Heart attacks among people with and without health insurance') +
        ylab('Total') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r heart attacks based on high blood pressure}
#Visualize proportion of respondents with heart attacks by high_bp
ggplot(brfss_complete, aes(x = high_bp, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Heart attacks among people with and without high blood pressure') +
        ylab('Total') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r sleep time and heart attack data}
#Filter out extreme sleep numbers to get ready to visualize
sleep_heartattack_df <- brfss_complete %>% select(sleep_time, heart_attack) %>%
   filter(as.integer(sleep_time)<14) %>%  filter(as.integer(sleep_time)>3) 
```

```{r sleep time and heart attack graph}
#Visualize sleep_time and heart_attack
ggplot(sleep_heartattack_df, aes(x = sleep_time, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Heart attacks among levels of sleep') +
        ylab('Total') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```
        
### Sarah EDA
Sarah's EDA, compare to heart attack:
- Time since cholesterols check and heart attack?\
- Smoking and heart attacks?\
- Angina and heart attack?\
- Depression and heart attack?\
- Kidney disease and heart attack?\
- Age

Cholesterol check
```{r}
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent = ifelse(time_since_cholcheck == "Within past year", "Recent", "Not Recent"))
  
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent_yes = ifelse(cholchkrecent == "Recent", 1, 0))
  
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent_no = ifelse(cholchkrecent == "Not Recent", 1, 0))



T_cholchkrecent <- brfss_complete %>%
  filter(told_high_chol != "No", !is.na(time_since_cholcheck), !is.na(age5yr_bucket)) %>%
  group_by(age5yr_bucket) %>% 
  summarise(cholchkratio = sum(cholchkrecent_yes) / sum(cholchkrecent_no), cholchkpercent = sum(cholchkrecent_yes) / (sum(cholchkrecent_no) + sum(cholchkrecent_yes)), cholchkcount = n())



#ratio of those checked
ggplot(data = T_cholchkrecent, aes(x = age5yr_bucket, y = cholchkratio)) +
       geom_point() +
       theme(axis.text.x = element_text(angle = 60, hjust = 1))

#percent of those checked
ggplot(data = T_cholchkrecent, aes(x = age5yr_bucket, y = cholchkpercent)) +
       geom_point() +
       theme(axis.text.x = element_text(angle = 60, hjust = 1))

T_cholchkrecent %>%
  group_by(age5yr_bucket) %>%
  summarise(cholchkratio, cholchkpercent, cholchkcount)
```

```{r}

ggplot(brfss_complete, aes(x=age5yr_bucket, y=time_since_cholcheck, color = heart_attack)) +
  geom_point(size=2) +
  geom_jitter()
```



Smoking and heart attack - TRY BAR CHART
```{r}
ggplot(brfss_complete, aes(x = smoke_100)) + geom_bar(aes(fill = heart_attack))

```

Angina - INTERESTING DISTRIBUTION, EXPLORE THIS FURTHER - BY AGE? BY GENDER?
```{r}
ggplot(brfss_complete, aes(x = angina)) + geom_bar(aes(fill = heart_attack))

```

Depression
```{r}
ggplot(brfss_complete, aes(x = depression)) + geom_bar(aes(fill = heart_attack))

```

Kidney disease- maybe look at kidney disease vs heart attack at different age brackets AND between genders, side by side bar chart
```{r}
ggplot(brfss_complete, aes(x = kidney_disease)) + geom_bar(aes(fill = heart_attack))

```

### Jesse EDA
- Smoking and heart attack?
- Smokeless tababcco and heart attack?
- Alcohol and heart attack?
- Fruit and heart attack?
- Veggies and heart attack?
- Exercise and heart attack?

```{r}
ggplot(brfss_complete, aes(x = smoke_100, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Number of participants to have smoked at least 100 cigarettes') +
        ylab('Total') +
        xlab('') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```        
        
```{r}
ggplot(brfss_complete, aes(x = smokeless_tabac, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Number of participants to have used smokeless tobacco products') +
        ylab('Total') +
        xlab('') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r}
hist(brfss_complete$alc_past_30,
     main = 'Frequency of alcohol consumption in 2013 among participants',
     xlab = 'Alcohol Consumption',
     col = magma(3),
     breaks=seq(0,250,by=50))
```

```{r}
hist(brfss_complete$fruit_freq,
     main = 'Frequency of fruit consumption in 2013 among participants',
     xlab = 'Fruit Consumption',
     col = viridis(7),
     breaks=seq(0,400,by=50))
```

```{r}
hist(brfss_complete$green_veg_freq,
     main = 'Frequency of dark green vegetable consumption in 2013 among participants',
     xlab = 'Dark Green Vegetable Consumption',
     col = cividis(3))
```

```{r}
ggplot(brfss_complete, aes(x = exercise_30, fill = age5yr_bucket)) +
        geom_bar() + 
        ggtitle('Number of participants to have exercised in the past 30 days') +
        ylab('Total') +
        xlab('') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
        scale_fill_manual(values=plasma(13))
```

### Ese EDA
- Diabetes and heart attack?
- Employment status and heart attack?
- Income and heart attack? 
- Weight and heart attack?
- Sex and heart attack?

```{r}
ggplot(brfss_complete, aes(x = diabetes, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Number of participants that heart attack among diabetes patients') +
        ylab('Total') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r}
ggplot(brfss_complete, aes(x = employment_status, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Participants with heart attack by employment status ') +
        ylab('Total') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r}
ggplot(brfss_complete, aes(x = income, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Participants with heart attack by Income distribution') +
        ylab('Total') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```

```{r}
brfss_complete$weight = as.numeric(brfss_complete$weight)
hist(brfss_complete$weight,
     main = 'Weight Distribution of Participants',
     xlab = 'Weight (in Pounds)',
     col = brfss_complete$heart_attack)
```

```{r}
ggplot(brfss_complete, aes(x = sex, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Participants with heart attack by sex') +
        ylab('Total') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## Hypothesis Tests

### Chi-Squared Tests of Independence

```{r}
#making contingency table for the categorical variable and heart_attack
contable1 = table(brfss_complete$gen_health, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest1 = chisq.test(contable1)
chi_p_value1 <- format(chitest1$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable2 = table(brfss_complete$mental_health, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest2 = chisq.test(contable2)
chi_p_value2 <- format(chitest2$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable3 = table(brfss_complete$health_coverage, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest3 = chisq.test(contable3)
chi_p_value3 <- format(chitest3$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable4 = table(brfss_complete$high_bp, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest4 = chisq.test(contable4)
chi_p_value4 <- format(chitest4$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable5 = table(brfss_complete$time_since_cholcheck, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest5 = chisq.test(contable5)
chi_p_value5 <- format(chitest5$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable6 = table(brfss_complete$told_high_chol, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest6 = chisq.test(contable6)
chi_p_value6 <- format(chitest6$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable7 = table(brfss_complete$angina, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest7 = chisq.test(contable7)
chi_p_value7 <- format(chitest7$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable8 = table(brfss_complete$stroke, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest8 = chisq.test(contable8)
chi_p_value8 <- format(chitest8$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable9 = table(brfss_complete$asthma, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest9 = chisq.test(contable9)
chi_p_value9 <- format(chitest9$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable10 = table(brfss_complete$depression, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest10 = chisq.test(contable10)
chi_p_value10 <- format(chitest10$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable11 = table(brfss_complete$kidney_disease, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest11 = chisq.test(contable11)
chi_p_value11 <- format(chitest11$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable12 = table(brfss_complete$diabetes, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest12 = chisq.test(contable12)
chi_p_value12 <- format(chitest12$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable13 = table(brfss_complete$veteran, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest13 = chisq.test(contable13)
chi_p_value13 <- format(chitest13$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable14 = table(brfss_complete$marital_status, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest14 = chisq.test(contable14)
chi_p_value14 <- format(chitest14$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable15 = table(brfss_complete$education_level, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest15 = chisq.test(contable15)
chi_p_value15 <- format(chitest15$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable16 = table(brfss_complete$employment_status, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest16 = chisq.test(contable16)
chi_p_value16 <- format(chitest16$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable17 = table(brfss_complete$income, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest17 = chisq.test(contable17)
chi_p_value17 <- format(chitest17$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable18 = table(brfss_complete$sex, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest18 = chisq.test(contable18)
chi_p_value18 <- format(chitest18$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable19 = table(brfss_complete$difficulty_walk, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest19 = chisq.test(contable19)
chi_p_value19 <- format(chitest19$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable20 = table(brfss_complete$smoke_100, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest20 = chisq.test(contable20)
chi_p_value20 <- format(chitest20$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable21 = table(brfss_complete$smokeless_tabac, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest21 = chisq.test(contable21)
chi_p_value21 <- format(chitest21$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable22 = table(brfss_complete$exercise_30, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest22 = chisq.test(contable22)
chi_p_value22 <- format(chitest22$p.value, scientific = FALSE)
```

```{r}
#making contingency table for the categorical variable and heart_attack
contable23 = table(brfss_complete$age5yr_bucket, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest23 = chisq.test(contable23)
chi_p_value23 <- format(chitest23$p.value, digits = 3)
```

Null Hypothesis: The two variables are independent.

Alternative Hypothesis: The two variables are dependent.

Categorical Variable  | Chi-Test P-value < 0.05?     |Include?   | 
----------------------|------------------------------|-----------|
gen_health            | `r chitest1$p.value < 0.05`  | Yes       | 
mental_health         | `r chitest2$p.value < 0.05`  | Yes       |  
health_coverage       | `r chitest3$p.value < 0.05`  | Yes       |  
high_bp               | `r chitest4$p.value < 0.05`  | Yes       |
time_since_cholcheck  | `r chitest5$p.value < 0.05`  | Yes       |
told_high_chol        | `r chitest6$p.value < 0.05`  | Yes       |
angina                | `r chitest7$p.value < 0.05`  | Yes       |
stroke                | `r chitest8$p.value < 0.05`  | Yes       |
ashtma                | `r chitest9$p.value < 0.05`  | Yes       |
depression            | `r chitest10$p.value < 0.05` | Yes       |
kidney_disease        | `r chitest11$p.value < 0.05` | Yes       |
diabetes              | `r chitest12$p.value < 0.05` | Yes       |
veteran               | `r chitest13$p.value < 0.05` | Yes       |
marital_status        | `r chitest14$p.value < 0.05` | Yes       |
education_level       | `r chitest15$p.value < 0.05` | Yes       |
employment_status     | `r chitest16$p.value < 0.05` | Yes       |
income                | `r chitest17$p.value < 0.05` | Yes       |
sex                   | `r chitest18$p.value < 0.05` | Yes       |
difficulty_walk       | `r chitest19$p.value < 0.05` | Yes       |
smoke_100             | `r chitest20$p.value < 0.05` | Yes       |
smokeless_tabac       | `r chitest21$p.value < 0.05` | No        |
exercise_30           | `r chitest22$p.value < 0.05` | Yes       |
age5yr_bucket         | `r chitest23$p.value < 0.05` | Yes       |

Based on the table above, we reject the null hypothesis for all variables except smokeless_tabac.  At a 0.05 threshold, all categorical variables have a dependent relationship with heart_attack except smokeless_tabac.  We will exclude smokeless_tabac in our models.  

## Modeling
  
### Logistic Regression

NEED TO NARROW DOWN TRAINING SET?
```{r}
# create test set and training set
set.seed(1)
heart_attack_sample <- sample(2, nrow(brfss_complete), replace=TRUE, prob=c(0.70, 0.30)) # ASK EDWIN ABOUT REPLACE = TRUE

# select columns x-y as predictor variables for test/train outputs
heart_attack_training <- brfss_complete[heart_attack_sample==1, c('heart_attack', 'state', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'time_since_cholcheck', 'angina', 'stroke', 'asthma', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'difficulty_walk', 'smoke_100', 'alc_past_30', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'age5yr_bucket')]

heart_attack_test <- brfss_complete[heart_attack_sample==2, c('heart_attack', 'state', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'time_since_cholcheck', 'angina', 'stroke', 'asthma', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'difficulty_walk', 'smoke_100', 'alc_past_30', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'age5yr_bucket')]

# create output variable for heart attack - NEED THIS?
# heart_attack.trainLabels <- brfss_complete[heart_attack_sample==1, 'heart_attack']
# heart_attack.testLabels <- brfss_complete[heart_attack_sample==2, 'heart_attack']

```

```{r}
# logistic regression - ONLY STATE OF FLORIDA??
# run chi squ test between categorical variables
# use forward selection method
heart_attack_logit1 <- glm(heart_attack ~ sex + sleep_time + age5yr_bucket, family="binomial", data = heart_attack_training)
  
# (state + gen_health + mental_health + health_coverage + sleep_time + high_bp + time_since_cholcheck + angina + stroke + asthma + depression + kidney_disease + diabetes + veteran + marital_status + education_level + employment_status + income + weight + sex + difficulty_walk + smoke_100 + alc_past_30 + fruit_freq + green_veg_freq + exercise_30 + age5yr_bucket, family="binomial", data = heart_attack_training)

summary(heart_attack_logit1)
```