---
title: "DATS6101 Group Project: Predicting Strokes"
author: "Jonathan Giguere, Jesse Borg, Ese Emuraye, Sarah Gates"
date: "December 11th 2019"
output: 
   html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

### Load packages

```{r load-packages, message = FALSE}
loadPkg('ggplot2')
loadPkg('dplyr')
```

### Load data

```{r load-data}
load("brfss2013.RData")
```

### Cleaning the data

```{r keeps}
# Choose which variables we want to keep

keeps <- c('X_state', 'imonth', 'genhlth', 'menthlth', 'hlthpln1', 'sleptim1', 'bphigh4', 'bpmeds', 'cholchk','toldhi2', 'cvdinfr4', 'cvdcrhd4', 'cvdstrk3', 'asthma3', 'asthnow', 'addepev2', 'chckidny', 'diabete3', 'veteran3', 'marital', 'educa', 'employ1', 'income2', 'weight2', 'sex', 'pregnant', 'diffwalk', 'smoke100', 'smokday2', 'stopsmk2', 'usenow3', 'alcday5', 'avedrnk2', 'drnk3ge5', 'fruit1', 'fvgreen', 'exerany2', 'prediab1', 'qlmentl2', 'qlstres2', 'drvisits', 'ssbsugar', 'cvdasprn', 'scntwrk1','X_ageg5yr')

brfss_reduced <- brfss2013[keeps]
```

```{r keeps}
# Rename the variables to be more descriptive

new_names <- c('state', 'month', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'bp_meds', 'time_since_cholcheck','told_high_chol', 'heart_attack', 'angina', 'stroke', 'asthma', 'has_asthma_now', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'pregnant', 'difficulty_walk', 'smoke_100', 'freq_smoke', 'stop_smoke_year', 'smokeless_tabac', 'alc_past_30', 'alc_perday_30', 'binge_alc', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'prediabetes', 'depressed_30', 'anxious_30', 'dr_visits_year', 'soda_freq', 'aspirin_daily', 'work_hours_week','age5yr_bucket')

names(brfss_reduced) <- new_names
```

```{r check for NAs}
# Check to see how many missing values for each column

missing_vals <- colSums(is.na(brfss_reduced))
sort(missing_vals)
```

```{r}
# Drop the variables which have too many missing values

brfss_reduced_drops <- subset(brfss_reduced, select = -c(dr_visits_year, prediabetes, binge_alc, alc_perday_30, freq_smoke, bp_meds, aspirin_daily, soda_freq, pregnant, stop_smoke_year, has_asthma_now, work_hours_week, depressed_30, anxious_30))
```

```{r}
# Remove the rows which have na values in any of the remaining variables

brfss_complete <- na.omit(brfss_reduced_drops)
```

```{r}
# Inspect the dataframe

str(brfss_complete)
```

For the factors, we want to change some of the variable factors to be more intuitive/descriptive, start by inspecting the levels for the factors we might want to change.

```{r}
# Check the levels for some varibles

levels(brfss_complete$high_bp)
levels(brfss_complete$employment_status)
levels(brfss_complete$smokeless_tabac)
levels(brfss_complete$diabetes)
```
Change the yes/no factors to 1/0 for model purposes and change other factors in the variables above to make them better grouped.

```{r}
# Change the factor levels for the variables wwe feel would be better off changed

levels(brfss_complete$health_coverage)[levels(brfss_complete$health_coverage) == 'Yes'] <- 1
levels(brfss_complete$health_coverage)[levels(brfss_complete$health_coverage) == 'No'] <- 0

levels(brfss_complete$heart_attack)[levels(brfss_complete$heart_attack) == 'Yes'] <- 1
levels(brfss_complete$heart_attack)[levels(brfss_complete$heart_attack) == 'No'] <- 0

levels(brfss_complete$angina)[levels(brfss_complete$angina) == 'Yes'] <- 1
levels(brfss_complete$angina)[levels(brfss_complete$angina) == 'No'] <- 0

levels(brfss_complete$stroke)[levels(brfss_complete$stroke) == 'Yes'] <- 1
levels(brfss_complete$stroke)[levels(brfss_complete$stroke) == 'No'] <- 0

levels(brfss_complete$asthma)[levels(brfss_complete$asthma) == 'Yes'] <- 1
levels(brfss_complete$asthma)[levels(brfss_complete$asthma) == 'No'] <- 0

levels(brfss_complete$depression)[levels(brfss_complete$depression) == 'Yes'] <- 1
levels(brfss_complete$depression)[levels(brfss_complete$depression) == 'No'] <- 0

levels(brfss_complete$kidney_disease)[levels(brfss_complete$kidney_disease) == 'Yes'] <- 1
levels(brfss_complete$kidney_disease)[levels(brfss_complete$kidney_disease) == 'No'] <- 0

levels(brfss_complete$veteran)[levels(brfss_complete$veteran) == 'Yes'] <- 1
levels(brfss_complete$veteran)[levels(brfss_complete$veteran) == 'No'] <- 0

levels(brfss_complete$difficulty_walk)[levels(brfss_complete$difficulty_walk) == 'Yes'] <- 1
levels(brfss_complete$difficulty_walk)[levels(brfss_complete$difficulty_walk) == 'No'] <- 0

levels(brfss_complete$smoke_100)[levels(brfss_complete$smoke_100) == 'Yes'] <- 1
levels(brfss_complete$smoke_100)[levels(brfss_complete$smoke_100) == 'No'] <- 0

levels(brfss_complete$exercise_30)[levels(brfss_complete$exercise_30) == 'Yes'] <- 1
levels(brfss_complete$exercise_30)[levels(brfss_complete$exercise_30) == 'No'] <- 0

levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == "Yes, but female told only during pregnancy"] <- 1
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == "No, pre-diabetes or borderline diabetes"  ] <- 0
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == 'Yes'] <- 1
levels(brfss_complete$diabetes)[levels(brfss_complete$diabetes) == 'No'] <- 0

levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == "Yes, but female told only during pregnancy"] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == "Told borderline or pre-hypertensive" ] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == 'Yes'] <- 1
levels(brfss_complete$high_bp)[levels(brfss_complete$high_bp) == 'No'] <- 0

levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Employed for wages" ] <- 'Employed'
levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Out of work for 1 year or more" ] <- 'Unemployed'
levels(brfss_complete$employment_status)[levels(brfss_complete$employment_status) == "Out of work for less than 1 year" ] <- 'Unemployed'

levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Every day" ] <- 1
levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Some days" ] <- 1
levels(brfss_complete$smokeless_tabac)[levels(brfss_complete$smokeless_tabac) == "Not at all" ] <- 0

```

```{r}
# Inspect the structure of the dataframe with changed variables

str(brfss_complete)
```


### EDA

Sarah's EDA, compare to heart attack:
- Time since cholesterols check and heart attack?\
- Smoking and heart attacks?\
- Angina and heart attack?\
- Depression and heart attack?\
- Kidney disease and heart attack?\
- Age

Cholesterol check
```{r}
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent = ifelse(time_since_cholcheck == "Within past year", "Recent", "Not Recent"))
  
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent_yes = ifelse(cholchkrecent == "Recent", 1, 0))
  
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent_no = ifelse(cholchkrecent == "Not Recent", 1, 0))



T_cholchkrecent <- brfss_complete %>%
  filter(told_high_chol != "No", !is.na(time_since_cholcheck), !is.na(age5yr_bucket)) %>%
  group_by(age5yr_bucket) %>% 
  summarise(cholchkratio = sum(cholchkrecent_yes) / sum(cholchkrecent_no), cholchkpercent = sum(cholchkrecent_yes) / (sum(cholchkrecent_no) + sum(cholchkrecent_yes)), cholchkcount = n())



#ratio of those checked
ggplot(data = T_cholchkrecent, aes(x = age5yr_bucket, y = cholchkratio)) +
       geom_point() +
       theme(axis.text.x = element_text(angle = 60, hjust = 1))

#percent of those checked
ggplot(data = T_cholchkrecent, aes(x = age5yr_bucket, y = cholchkpercent)) +
       geom_point() +
       theme(axis.text.x = element_text(angle = 60, hjust = 1))

T_cholchkrecent %>%
  group_by(age5yr_bucket) %>%
  summarise(cholchkratio, cholchkpercent, cholchkcount)
```

```{r}

ggplot(brfss_complete, aes(x=age5yr_bucket, y=time_since_cholcheck, color = heart_attack)) +
  geom_point(size=2) +
  geom_jitter()
```


Smoking and heart attack - TRY BAR CHART
```{r}
ggplot(brfss_complete, aes(x = heart_attack)) + geom_bar(aes(fill = smoke_100))

```

Angina - INTERESTING DISTRIBUTION, EXPLORE THIS FURTHER - BY AGE? BY GENDER?
```{r}
ggplot(brfss_complete, aes(x = heart_attack)) + geom_bar(aes(fill = angina))

```

Depression
```{r}
ggplot(brfss_complete, aes(x = heart_attack)) + geom_bar(aes(fill = depression))

```

Kidney disease- maybe look at kidney disease vs heart attack at different age brackets AND between genders, side by side bar chart
```{r}
ggplot(brfss_complete, aes(x = heart_attack)) + geom_bar(aes(fill = kidney_disease))

```


Perform correlation plot

## Section 3
  
### 3.1: Logistic Regression


NEED TO NARROW DOWN TRAINING SET?
```{r}
# create test set and training set
set.seed(1)
heart_attack_sample <- sample(2, nrow(brfss_complete), replace=TRUE, prob=c(0.70, 0.30)) # ASK EDWIN ABOUT REPLACE = TRUE

# select columns x-y as predictor variables for test/train outputs
heart_attack_training <- brfss_complete[heart_attack_sample==1, c('heart_attack', 'state', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'time_since_cholcheck', 'angina', 'stroke', 'asthma', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'difficulty_walk', 'smoke_100', 'alc_past_30', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'age5yr_bucket')]

heart_attack_test <- brfss_complete[heart_attack_sample==2, c('heart_attack', 'state', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'time_since_cholcheck', 'angina', 'stroke', 'asthma', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'difficulty_walk', 'smoke_100', 'alc_past_30', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'age5yr_bucket')]

# create output variable for heart attack - NEED THIS?
# heart_attack.trainLabels <- brfss_complete[heart_attack_sample==1, 'heart_attack']
# heart_attack.testLabels <- brfss_complete[heart_attack_sample==2, 'heart_attack']

```

```{r}
# logistic regression - ONLY STATE OF FLORIDA??
# run chi squ test between categorical variables
# use forward selection method
heart_attack_logit1 <- glm(heart_attack ~ sex + sleep_time + age5yr_bucket, family="binomial", data = heart_attack_training)
  
# (state + gen_health + mental_health + health_coverage + sleep_time + high_bp + time_since_cholcheck + angina + stroke + asthma + depression + kidney_disease + diabetes + veteran + marital_status + education_level + employment_status + income + weight + sex + difficulty_walk + smoke_100 + alc_past_30 + fruit_freq + green_veg_freq + exercise_30 + age5yr_bucket, family="binomial", data = heart_attack_training)

summary(heart_attack_logit1)
```

