---
title: "DATS6101 Group Project: Predicting Heart Attacks - EDA Code"
author: "Jonathan Giguere, Jesse Borg, Ese Emuraye, Sarah Gates"
date: "December 11th 2019"
output:
  html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

Start by loading the packages we know will be needed for the presentation. Then load the brfss 2013 data which is as an RData file as opposed to a csv.

```{r load-packages, message = FALSE}
# Load the packages using the loadPkg function

loadPkg('ggplot2')
loadPkg('dplyr')
loadPkg('usmap')
loadPkg('viridis')
loadPkg('ResourceSelection')
loadPkg('pROC')
loadPkg('pscl')
loadPkg('caret')
loadPkg('car')
```

```{r load-data}
# Load the RData dataset
load("brfss_complete.RData")
```

## Graphs

Plot showing the number of participants by state:

```{r states and participants data}
#Make dataframe showing number of survey particpants by state
participants_map_df <- participants_map_df[-c(52, 53), ]
participants_map_df <-  brfss_complete %>% group_by(state) %>% summarise(n = n())
```

```{r states and respondents}
#Visualize map of number of participants per state
plot_usmap(data = participants_map_df, values = "n", color = "white") + 
  scale_fill_continuous(name = "Survey Participants by State", label = scales::comma) + 
  labs(title = 'Number of Survey Participants per State') +
  theme(legend.position = "right") 
```

Plot percentage of heart attacks to participants by state

```{r states and heart attack data}
#Making dataframe for visualizing heart attacks per state
#Notice that in order to add up the number of heart attacks, i convert the factor to a string and then an integer
state_map_df <- brfss_complete %>% select(state, heart_attack) %>% mutate(state = as.character(state)) %>% group_by(state) %>% summarise(sum_ = sum(as.numeric(as.character(heart_attack)))) %>% filter(state != c('District of Columbia', 'Puerto Rico', 'Guam'))

# Create the percentage of heart attacks to participants
ratios <- (state_map_df$sum_ * 100)/ participants_map_df$n
ratio_df <- data.frame(state_map_df$state, ratios)
names(ratio_df) <- c("state", 'ratio')
```

```{r states and heart map}
#Map showing number of heart attacks per state as a percentage of participants

plot_usmap(data = ratio_df, values = "ratio", color = "white") + 
  scale_fill_continuous(name = "Heart Attack Ratio", label = scales::comma) + 
  labs(title = 'Ratio of Heart Attacks to Participants per State') +
  theme(legend.position = "right")
```

Graph of how the porportion of heart attacks vary depending on the general health of the participants. 

```{r heart attacks for different gen_health levels}
#Visualize proportion of respondents with heart attacks by gen_health
ggplot(brfss_complete, aes(x = gen_health, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Heart attack among levels of general health') +
        ylab('Total') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Graphs showing the ratio and percentage of cholesterol checks by 5 year age bucket:

First, classify Recent and Not Recent under last cholesterol check as whether the check occurred in the last year or not. Then create a subset of only individuals told that they have high cholesterol.

```{r}
# Change the responses from 'within past year' to 'recent' if yes and 'not recent' if no 
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent = ifelse(time_since_cholcheck == "Within past year", "Recent", "Not Recent"))
  
# Create a column for those who responded 'recent',
# Change the responses from 'recent' to 1 if yes and 0 if no
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent_yes = ifelse(cholchkrecent == "Recent", 1, 0))

# Create a column for those who responded 'not recent'
# Change the responses from 'not recent' to 1 if yes and 0 if no  
brfss_complete <- brfss_complete %>%
  mutate(cholchkrecent_no = ifelse(cholchkrecent == "Not Recent", 1, 0))


# Create a subsetted dataframe to only have those which had a recent cholesterol check,
# then create summart columns  
T_cholchkrecent <- brfss_complete %>%
  filter(told_high_chol != "No", !is.na(time_since_cholcheck), !is.na(age5yr_bucket)) %>%
  group_by(age5yr_bucket) %>% 
  summarise(cholchkratio = sum(cholchkrecent_yes) / sum(cholchkrecent_no), cholchkpercent = sum(cholchkrecent_yes) / (sum(cholchkrecent_no) + sum(cholchkrecent_yes)), cholchkcount = n()) 
# cholchkratio: ratio of cholchkrecent yes to no
# cholchkpercent: percent of cholchkrecent yes over total people with high cholesterol

# Plots
# Plot the ratio of those checked
ggplot(data = T_cholchkrecent, aes(x = age5yr_bucket, y = cholchkratio, color = cholchkratio, size = 1)) +
       geom_point(show.legend = FALSE) +
       ggtitle('Ratio of Cholesterol Checks by 5 Year Age Bucket') +
       ylab('Cholesterol Check Ratio') +
       xlab('Age (5 Year Bucket)') +
       theme_bw() +
       theme(plot.title = element_text(hjust = 0.5)) +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, color = 'black')) +
       theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())


# Plot the percentage of those checked
ggplot(data = T_cholchkrecent, aes(x = age5yr_bucket, y = cholchkpercent, color = cholchkpercent, size = 1)) +
       geom_point(show.legend = FALSE) +
       ggtitle('Percentage of Cholesterol Checks by 5 Year Age Bucket') +
       ylab('Cholesterol Check Percentage') +
       xlab('Age (5 Year Bucket)') +
       theme_bw() +
       theme(plot.title = element_text(hjust = 0.5)) +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, color = 'black')) +
       theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# Group by the 5 year buckets and show the ratio, percentages and count 
T_cholchkrecent %>%
  group_by(age5yr_bucket) %>%
  summarise(cholchkratio, cholchkpercent, cholchkcount)
```

Graph showing the ratio of heart attacks vary with the participants by their age and the time since their last cholesterol check.  

```{r}
# Visualize the ratio of heart attacks by age and time since last cholesterol check
ggplot(brfss_complete, aes(x=age5yr_bucket, y=time_since_cholcheck, color = heart_attack)) +
  geom_point(size=2) +
  geom_jitter() +
  ggtitle('Ratio of Heart Attacks by Time Since Cholesterol Check and 5 Year Age') +
  ylab('Time Since Cholesterol Check') +
  xlab('Age (5 Year Bucket)') +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, color = 'black')) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Graph showing the ratio of heart attacks depending on whether the participants have had an angina or not.

```{r}
# Visualize the ratio of heart attacks to angina
ggplot(brfss_complete, aes(x = angina)) + geom_bar(aes(fill = heart_attack)) +
       ggtitle('Ratio of Heart Attacks if Participant Has Suffered from Angina or Not') +
       ylab('Count') +
       xlab('Suffered Angina?') +
       theme_bw() +
       theme(plot.title = element_text(hjust = 0.5)) +
       theme(axis.text.x = element_text(size = 11, color = 'black')) +
       theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Graph showing the number of participants to have exercised within the past 30 days by 5 year age bucket.
  
```{r}
# Visualize if exercised in past 30 days by 5 year age bucket
ggplot(brfss_complete, aes(x = exercise_30, fill = age5yr_bucket)) +
        geom_bar() + 
        ggtitle('Number of participants to have exercised in the past 30 days') +
        ylab('Total') +
        xlab('Exercised in the Past 30 Days?') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
        scale_fill_manual(values=plasma(13))
```

Graph showing the ratio of heart attacks depending on whether participants have exercised in the past 30 days products or not.
  
```{r}
# Visualize ratio of heart attacks if exercised in past 30 days
ggplot(brfss_complete, aes(x = exercise_30, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Ratio of Heart Attacks if Participant Has Exercised in the Past 30 Days or Not') +
        ylab('Total') +
        xlab('Exercised in the Past 30 Days?') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Graph showing the ratio of heart attacks depending on the participants' employment status.

```{r}
# Visualize heart attack ratio to employment
ggplot(brfss_complete, aes(x = employment_status, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Number of Participants With Heart Attacks by Employment Status ') +
        ylab('Total') +
        xlab('Employment Status') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 10, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Graph showing the ratio of heart attacks depending on the participants' annual income.

```{r}
# Visualize heart attack ratio to annual income
ggplot(brfss_complete, aes(x = income, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Number of Participants With Heart Attacks by Annual Income') +
        ylab('Total') +
        xlab('Annual Income') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 10, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```

Graph showing the ratio of heart attacks depending on the participants' gender.

```{r}
# Visualize heart attack ratio to gender
ggplot(brfss_complete, aes(x = sex, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Number pf Participants With Heart Attacks by Gender') +
        ylab('Total') +
        xlab('Gender') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Graph showing how the porportion of heart attacks vary depending on whether the participants have high blood pressure or not.

```{r heart attacks based on high blood pressure}
#Visualize proportion of respondents with heart attacks by high_bp
ggplot(brfss_complete, aes(x = high_bp, fill = heart_attack)) +
        geom_bar() + 
        ggtitle('Heart attacks among people with and without high blood pressure') +
        ylab('Total') +
        scale_x_discrete(labels= c('Yes', 'No')) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(size = 11, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## Hypothesis Testing

Before the models are built, some hypothesis testing will be carried out to see which variables are dependent or independent from the heart attacks.

### Chi-Squared Tests of Independence

Start off by carrying out chi-squared test of independence for the different categorical variables.

<br></br>

Chi-Squared Test for General Health And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable1 = table(brfss_complete$gen_health, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest1 = chisq.test(contable1)
chi_p_value1 <- format(chitest1$p.value, scientific = FALSE)
```

Chi-Squared Test for Mental Health And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable2 = table(brfss_complete$mental_health, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest2 = chisq.test(contable2)
chi_p_value2 <- format(chitest2$p.value, scientific = FALSE)
```

Chi-Squared Test for Health Coverage And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable3 = table(brfss_complete$health_coverage, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest3 = chisq.test(contable3)
chi_p_value3 <- format(chitest3$p.value, scientific = FALSE)
```

Chi-Squared Test for High Blood Pressure And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable4 = table(brfss_complete$high_bp, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest4 = chisq.test(contable4)
chi_p_value4 <- format(chitest4$p.value, scientific = FALSE)
```

Chi-Squared Test for Time Since Last Cholesterol Check And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable5 = table(brfss_complete$time_since_cholcheck, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest5 = chisq.test(contable5)
chi_p_value5 <- format(chitest5$p.value, scientific = FALSE)
```

Chi-Squared Test for Told If Cholesterol Was High And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable6 = table(brfss_complete$told_high_chol, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest6 = chisq.test(contable6)
chi_p_value6 <- format(chitest6$p.value, scientific = FALSE)
```

Chi-Squared Test for Angina And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable7 = table(brfss_complete$angina, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest7 = chisq.test(contable7)
chi_p_value7 <- format(chitest7$p.value, scientific = FALSE)
```

Chi-Squared Test for Stroke And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable8 = table(brfss_complete$stroke, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest8 = chisq.test(contable8)
chi_p_value8 <- format(chitest8$p.value, scientific = FALSE)
```

Chi-Squared Test for Asthma And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable9 = table(brfss_complete$asthma, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest9 = chisq.test(contable9)
chi_p_value9 <- format(chitest9$p.value, scientific = FALSE)
```

Chi-Squared Test for Depression And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable10 = table(brfss_complete$depression, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest10 = chisq.test(contable10)
chi_p_value10 <- format(chitest10$p.value, scientific = FALSE)
```

Chi-Squared Test for Kidney Disease And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable11 = table(brfss_complete$kidney_disease, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest11 = chisq.test(contable11)
chi_p_value11 <- format(chitest11$p.value, scientific = FALSE)
```

Chi-Squared Test for Diabetes And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable12 = table(brfss_complete$diabetes, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest12 = chisq.test(contable12)
chi_p_value12 <- format(chitest12$p.value, scientific = FALSE)
```

Chi-Squared Test for Veteran Status And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable13 = table(brfss_complete$veteran, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest13 = chisq.test(contable13)
chi_p_value13 <- format(chitest13$p.value, scientific = FALSE)
```

Chi-Squared Test for Marital Status And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable14 = table(brfss_complete$marital_status, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest14 = chisq.test(contable14)
chi_p_value14 <- format(chitest14$p.value, scientific = FALSE)
```

Chi-Squared Test for Education Level And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable15 = table(brfss_complete$education_level, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest15 = chisq.test(contable15)
chi_p_value15 <- format(chitest15$p.value, scientific = FALSE)
```

Chi-Squared Test for Employment Status And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable16 = table(brfss_complete$employment_status, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest16 = chisq.test(contable16)
chi_p_value16 <- format(chitest16$p.value, scientific = FALSE)
```

Chi-Squared Test for Annual Income And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable17 = table(brfss_complete$income, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest17 = chisq.test(contable17)
chi_p_value17 <- format(chitest17$p.value, scientific = FALSE)
```

Chi-Squared Test for Gender And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable18 = table(brfss_complete$sex, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest18 = chisq.test(contable18)
chi_p_value18 <- format(chitest18$p.value, scientific = FALSE)
```

Chi-Squared Test for Difficulty Walking And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable19 = table(brfss_complete$difficulty_walk, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest19 = chisq.test(contable19)
chi_p_value19 <- format(chitest19$p.value, scientific = FALSE)
```

Chi-Squared Test for Smoked at Least 100 Cigarettes And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable20 = table(brfss_complete$smoke_100, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest20 = chisq.test(contable20)
chi_p_value20 <- format(chitest20$p.value, scientific = FALSE)
```

Chi-Squared Test for Use of Smokeless Tobacco Products And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable21 = table(brfss_complete$smokeless_tabac, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest21 = chisq.test(contable21)
chi_p_value21 <- format(chitest21$p.value, scientific = FALSE)
```

Chi-Squared Test for Exercise in the Past 30 Days And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable22 = table(brfss_complete$exercise_30, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest22 = chisq.test(contable22)
chi_p_value22 <- format(chitest22$p.value, scientific = FALSE)
```

Chi-Squared Test for Age (5 Year Bucket) And Heart Attack

```{r}
#making contingency table for the categorical variable and heart_attack
contable23 = table(brfss_complete$age5yr_bucket, brfss_complete$heart_attack)

#throw contingency table into chitest function and get p-value from test
chitest23 = chisq.test(contable23)
chi_p_value23 <- format(chitest23$p.value, digits = 3)
```

<br></br>

After all of the Chi-Squared Tests of Independence are carried out, create a table to summarise the P-values for each test and also whether to include them in our modelling or not.

<br></br>

Null Hypothesis: The two variables are independent.

Alternative Hypothesis: The two variables are dependent.

<br></br>

Categorical Variable  | Chi-Test P-value < 0.05?     |Include?   | 
----------------------|------------------------------|-----------|
gen_health            | `r chitest1$p.value < 0.05`  | Yes       | 
mental_health         | `r chitest2$p.value < 0.05`  | Yes       |  
health_coverage       | `r chitest3$p.value < 0.05`  | Yes       |  
high_bp               | `r chitest4$p.value < 0.05`  | Yes       |
time_since_cholcheck  | `r chitest5$p.value < 0.05`  | Yes       |
told_high_chol        | `r chitest6$p.value < 0.05`  | Yes       |
angina                | `r chitest7$p.value < 0.05`  | Yes       |
stroke                | `r chitest8$p.value < 0.05`  | Yes       |
ashtma                | `r chitest9$p.value < 0.05`  | Yes       |
depression            | `r chitest10$p.value < 0.05` | Yes       |
kidney_disease        | `r chitest11$p.value < 0.05` | Yes       |
diabetes              | `r chitest12$p.value < 0.05` | Yes       |
veteran               | `r chitest13$p.value < 0.05` | Yes       |
marital_status        | `r chitest14$p.value < 0.05` | Yes       |
education_level       | `r chitest15$p.value < 0.05` | Yes       |
employment_status     | `r chitest16$p.value < 0.05` | Yes       |
income                | `r chitest17$p.value < 0.05` | Yes       |
sex                   | `r chitest18$p.value < 0.05` | Yes       |
difficulty_walk       | `r chitest19$p.value < 0.05` | Yes       |
smoke_100             | `r chitest20$p.value < 0.05` | Yes       |
smokeless_tabac       | `r chitest21$p.value < 0.05` | No        |
exercise_30           | `r chitest22$p.value < 0.05` | Yes       |
age5yr_bucket         | `r chitest23$p.value < 0.05` | Yes       |

Based on the table above, we reject the null hypothesis for all variables except smokeless_tabac.  At a 0.05 threshold, all categorical variables have a dependent relationship with heart_attack except smokeless_tabac.  We will exclude smokeless_tabac in our models.

## Modeling

After carrying out the Chi-Squared Test of Indepence to determine which categorical variables to use in our models, its time to build the models. Start off by using classification trees.

### Getting Test and Training Sets

Before diving into the creation of our first classification tree, we noticed that the data is imbalanced.  This is a common occurence when working with medical data.  Only 6.1% of the participants had heart attacks in our cleaned up dataset. In order to balance our dataset for training and testing, we identified 19212 people with heart attacks and subset them.  Then, we randomly samples 19212 other participants from the dataframe that contained no heart attacks.  Finally, we put these two dataframes together to get a balanced dataset with a total of 38424 records.  


```{r}
# Create two subsets, one where heart attack occured (1) and one where heart attack didnt occur (0)
brfss_hrtattack <- subset(brfss_complete, heart_attack == '1')
brfss_no_hrtattack <- subset(brfss_complete, heart_attack == '0')

# Create ratios for the number of heart attacks occured and heart attacks didnt occur
ratio1 <- nrow(brfss_hrtattack)/nrow(brfss_complete)
ratio2 <- nrow(brfss_no_hrtattack)/nrow(brfss_complete)

# Set the seen and then take a sample of no heart attacks randomly to balance the data
set.seed(1)
brfss_no_hrtattack <- sample_n(brfss_no_hrtattack, nrow(brfss_hrtattack))

# Display the results
balanced_brfss = rbind(brfss_hrtattack, brfss_no_hrtattack)
nrow(balanced_brfss)
rownames(balanced_brfss) <- NULL
tail(balanced_brfss, 5)
```

Now that we have a balanced dataset, we split the data into training and test sets.  We opted to use 30% of the data for testing and 70% for training.

```{r}
# create test set and training set
set.seed(1)
balanced_hrt_attack_sample <- sample(2, nrow(balanced_brfss), replace=TRUE, prob=c(0.70, 0.30))

# select columns x-y as predictor variables for test/train outputs
bal_hrt_attack_training <- balanced_brfss[balanced_hrt_attack_sample==1, c('heart_attack', 'state', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'time_since_cholcheck', 'angina', 'stroke', 'asthma', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'difficulty_walk', 'smoke_100', 'alc_past_30', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'age5yr_bucket')]

bal_hrt_attack_test <- balanced_brfss[balanced_hrt_attack_sample==2, c('heart_attack', 'state', 'gen_health', 'mental_health', 'health_coverage', 'sleep_time', 'high_bp', 'time_since_cholcheck', 'angina', 'stroke', 'asthma', 'depression', 'kidney_disease', 'diabetes', 'veteran', 'marital_status', 'education_level', 'employment_status', 'income', 'weight', 'sex', 'difficulty_walk', 'smoke_100', 'alc_past_30', 'fruit_freq', 'green_veg_freq', 'exercise_30', 'age5yr_bucket')]

# create output variable for heart attack
#heart_attack.trainLabels <- brfss_complete[heart_attack_sample==1, 'heart_attack']
#heart_attack.testLabels <- brfss_complete[heart_attack_sample==2, 'heart_attack']
```

### Decision Trees

First, we construct a classification tree using all variables as predictors except smokeless_tabac (excluded as a result of chi-square tests), state, and month. We are using our balanced training data for creation of the tree.

```{r}
# LOoad the tree package
loadPkg('tree')

# Create a tree to predict heart attack using all the variables
heart_attack_fit2 <- tree(heart_attack ~ gen_health + mental_health + health_coverage + sleep_time + high_bp + time_since_cholcheck + angina + stroke + asthma + depression + kidney_disease + diabetes + veteran + marital_status + education_level + employment_status + income + weight + sex + difficulty_walk + smoke_100 + alc_past_30 + fruit_freq + green_veg_freq + exercise_30 + age5yr_bucket, bal_hrt_attack_training)

# Summarize the results
summary(heart_attack_fit2)

# Plot the results
plot(heart_attack_fit2, uniform = TRUE, main="Classification Tree")
text(heart_attack_fit2, use.n = TRUE, all = TRUE, cex=.8)
```

We can see that the variables in order of importance for determining heart attack are: angina, employment_status, gen_health, and high_bp.  Many of the predictor variables are ignored as part of the tree function because their contributions to the model are very insignificant

After creating the above tree, we want to assess how it classifies our training cases.  To do so, we will create a confusion matrix and assess metrics like accuracy, precision, and recall.

```{r, include=T}
# Create a matrix to predict the heart attacks using the training set
cm = confusionMatrix( predict(heart_attack_fit2, type = "class"), reference = bal_hrt_attack_training[, "heart_attack"] )

# Print the results
print('Overall: ')
cm$overall
print('Class: ')
cm$byClass
```

We now look at a confusion matrix to see how well the tree classifies our test cases.  We also have calculated accuracy below.  We can see that our first decision tree classified the test cases with 77.6% accuracy.  

```{r}
# Run the prediction for the heart attack test data
tree.pred = predict(heart_attack_fit2, bal_hrt_attack_test, type = "class")

# Create a table of the results
cmm = table(tree.pred, bal_hrt_attack_test$heart_attack)
cmm

# Show the accuracy
accuracy = (cmm[[1]] + cmm[[4]])/sum(cmm)
accuracy
```

Next we will perform some cross validation to decide if our tree would benefit from any pruning.

```{r}
# Run the cross validation of the tree, show the number of splits and the errors of each cross validation
cv.heart_attack = cv.tree(heart_attack_fit2, FUN=prune.misclass)
cv.heart_attack$size
cv.heart_attack$dev
```

Based on the cross validation, we can see that the optimal number of splits for the tree are either 5 or 4.  Because of these results, we will keep our original tree with four splits, as pruning the tree would provide no benefit.

Next we attempt to perform bagging.  This involves using all predictor variables to make a whole bunvh of trees.  Using more than one tree will hopefully increase model performance.

```{r}
# Load randomforest package
loadPkg('randomForest')

# Subset the training data to exclude the weight and state of each participant
bal_hrt_attack_training <- subset(bal_hrt_attack_training, select = -c(weight, state))

# Create a forest for the subsetted data
bag.heart_attack = randomForest(heart_attack~., data=bal_hrt_attack_training, mtry=26, importance= TRUE)
bag.heart_attack
```

We can see that the performance of the bagged tree performed with slightly better accuracy.

```{r}
# Run the prediction for the bagged tree
tree.pred_bagged = predict(bag.heart_attack, bal_hrt_attack_test, type = "class")

# Create a table for the results
cmm1 = table(tree.pred_bagged, bal_hrt_attack_test$heart_attack)
cmm1

# Calculate and show the accuracy of the results
accuracy1 = (cmm1[[1]] + cmm1[[4]])/sum(cmm1)
accuracy1
```

Next we will look at random forest and see if we can get even better accuracy.

```{r}
# Use the subsetted data from before without the weight and state,
# Create a forest using different parameters
randForest.heart_attack = randomForest(heart_attack~., data=bal_hrt_attack_training, mtry=2, importance= TRUE)
randForest.heart_attack
```

The confusion matrix and accuracy for the random forest is given below.

```{r}
# Run the predictions for the random forest above
tree.pred_randForest = predict(randForest.heart_attack, bal_hrt_attack_test, type = "class")

# Plot a table for the results
cmm2 = table(tree.pred_randForest, bal_hrt_attack_test$heart_attack)
cmm2

# Plot the accuracy
accuracy2 = (cmm2[[1]] + cmm2[[4]])/sum(cmm2)
accuracy2
```

### Logistic Regression

Key variables identified in trees in order of significance: agina, employment status, gen health, high bp

```{r}
# logistic regression
# run chi squ test between categorical variables
# use forward selection method
heart_attack_logit1 <- glm(heart_attack ~ angina + employment_status + gen_health + high_bp, family="binomial", data = bal_hrt_attack_training)

# all variables
# ( sex + sleep_time + age5yr_bucket+ state + gen_health + mental_health + health_coverage + sleep_time + high_bp + time_since_cholcheck + angina + stroke + asthma + depression + kidney_disease + diabetes + veteran + marital_status + education_level + employment_status + income + weight + sex + difficulty_walk + smoke_100 + alc_past_30 + fruit_freq + green_veg_freq + exercise_30 + age5yr_bucket, family="binomial", data = heart_attack_training)

summary(heart_attack_logit1)
```
  
Growth/decay factors for each explanatory variable (calculated as exponentials of the model coefficients):

```{r growthDecayFactors, results='markup', collapse=F}
# Calculate the exponentials
exp(coef(heart_attack_logit1))
```
  
Confidence intervals of each coefficient of the model:

```{r ConfInt, results='markup', collapse=F}
# CIs using standard errors
confint.default(heart_attack_logit1)
```
  
Hosmer and Lemeshow Goodness of Fit:

```{r HosmerLemeshow}
# Carry out Hosmer & Lemeshow GOF
heart_attackLogitHoslem = hoslem.test(bal_hrt_attack_training$heart_attack, fitted(heart_attack_logit1))
heart_attackLogitHoslem
```
  
According to the Hosmer and Lemeshow Goodness of Fit test using the training dataset, the p value is `r heart_attackLogitHoslem$p.value`. This is very small so it means that this model is a good fit. Although this test is not very meaningful for models with only factors, it continues to support the other analysis. 

McFadden test:

```{r McFadden}
# Carry out the McFadden Test
heart_attack_logit1_pr2 = pR2(heart_attack_logit1)
heart_attack_logit1_pr2
```

The McFadden value of this model is `r heart_attack_logit1_pr2['McFadden']`, which means that about 35% of the variations in y is explained by the variables in the model.  
  
ROC & AUC analysis:  
Use training and test datasets to evaluate the area under the curve (AUC) for each.  
  
Training:

```{r roc_auc}
# Carry out the probability calculation for the training set
# prob = predict(heart_attack_logit1, bal_hrt_attack_training, type = "response")
prob = plogis(predict(heart_attack_logit1, bal_hrt_attack_training))
bal_hrt_attack_training$prob=prob

# Create the roc and then calculate and plot the area under the curve
h <- roc(heart_attack~prob, data=bal_hrt_attack_training)
auc(h) # area-under-curve (prefer 0.8 or higher)
plot(h)
```

The the area-under-curve is `r auc(h)` for the training data, which is more than 0.8. This outcome agrees with the Hosmer and Lemeshow test that the model is considered a good fit. More importantly, we will now look at the test data.   
  
Test set: 

```{r roc_auc2}
# Carry out the probability calculation for the test set
# prob2 = predict(heart_attack_logit1, bal_hrt_attack_test, type = "response") # use model on test data
prob2 = plogis(predict(heart_attack_logit1, bal_hrt_attack_test))
bal_hrt_attack_test$prob=prob2

# Create the roc and then calculate and plot the area under the curve
j <- roc(heart_attack~prob, data=bal_hrt_attack_test)
auc(j) # area-under-curve (prefer 0.8 or higher)
plot(j)
```

The the area-under-curve is `r auc(j)` which is also above .8, meaning that this is a strong model.  
  
Create confusion matrix from logistic regression output  

```{r cm}
# VIF analysis
# vif(heart_attack_logit1)

# confusion matrix for test
cmlog = table(bal_hrt_attack_test$heart_attack, prob2 > .5)
cmlog

testSensitivity <- 1573 / (4187 + 1573)
testSpecificity <- 953 / (953 + 4803)
testPrecision <- 1573 / (1573 + 4803)
testAccuracy <- (1573 + 953) / (4187 + 1573 + 953 + 4803)

# F1 score
testF1 <- 2 * testPrecision * testSensitivity / (testPrecision + testSensitivity)
```
  
## Comparisons Between Models   
  
Statistic              | Logistic Regression | Classification Tree | Bagged Tree | Random Forrest |
-----------------------|---------------------|---------------------|-------------|----------------|
AUC                    | `r auc(j)`          |                     |             |                |
Precision              | `r testPrecision`   |                     |             |                |
Accuracy               | `r testAccuracy`    |                     |             |                |
Specificity            | `r testSpecificity` |                     |             |                | 
Sensitivity/Recall     | `r testSensitivity` |                     |             |                |  
F1 Score               | `r testF1`          |                     |             |                | 
